Methods paper for erosion

### 2017-08-30 docs/methods/erosion_methods.tex
worked on section 2 of the methods paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Formulation\label{s:formulation}} 
Let us first define the main variables used to model erosion.  We only
consider flows that are confined by a solid wall $\Gamma$ that encloses
$M$ eroding bodies.  The bodies are denoted as $\gamma_k$,
$k=1,\ldots,M$, and we write $\gamma = \gamma_1 \cup \cdots \cup
\gamma_M$.  Neglecting inertial forces, the dynamics of the fluid is
fully characterized by the position of the bodies $\xx_k(s,t) \in
\gamma_k$, where $s$ is the arclength and $t$ is time.  Given
$\xx_k(s,t)$, $k=1,\ldots,M$, derived variables include the fluid
velocity $\uu$, the pressure $p$, and the shear stress $\tau$.   On the
bounding wall, we prescribed a velocity $\UU(\xx,t)$.  Then, the
governing equations are
\begin{equation}
\label{eqn:erosionModel}
\begin{split}
  \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap &&\mbox{conservation
of momentum}\\
\grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap
&&\mbox{conservation of mass} \\
\uu = 0, &\hspace{20pt} \xx \in \gamma, \gap &&\mbox{no slip on the
bodies} \\
\uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap &&\mbox{outer wall
velocity} \\
\dot{\xx}(t) = -|\tau| \nn, &\hspace{20pt} \xx \in \gamma,
&&\mbox{erosion model},
\end{split}
\end{equation}
where $\nn$ is the outward unit normal vector which points away from the
fluid.  The outer wall is formed by rounding off the corners of $[-3,3]
\times [-1,1]$, and we enforce the Hagen-Poiseuille flow
\begin{align}
  \uu(\xx) = (1-y^2,0)
\end{align}
on $\Gamma$.  To minimize the boundary effects of the inflow and
outflow, we only place bodies in the center third, $[-1,1] \times
[-1,1]$, of the fluid domain.
\begin{figure}[htpb]
  \centering
  \input{figs/schematic.tikz}
  \caption{\label{fig:schematic} A schematic for the governing
    equations.  A no-slip boundary condition is imposed on each body
    $\gamma_k$ whose unit normal points outward relative to the
    geometry.  On the outer geometry $\Gamma$, a pipe flow is imposed.
    Finally, the bodies are eroded according to the shear stress applied
    by the Stokesian fluid.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Boundary Integral Equation Formulation} 
\label{sec:bies}
We use a quasi-static approximation.  That is, at each time step, we
freeze the geometry, solve the incompressible Stokes equations, and then
update the geometry according to the erosion model.  A flow diagram is
illustrated in Figure~\ref{fig:workflow}.

The incompressible Stokes equations are
\begin{alignat*}{3}
  \mu\Delta \uu &= \grad p, \qquad && \xx \in \Omega, \\
  \grad \cdot \uu &= 0,   && \xx \in \Gamma.
\end{alignat*}
The viscosity $\mu$ sets the time scale, and we will assume that
$\mu=1$.  The double-layer potential
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{\pi} \int_{\Gamma} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2} 
    \eeta \, ds_\yy, \quad \xx \in \Omega,
    \label{eqn:velocityDLP}
\end{align}
where $\rr = \xx - \yy$, $\rho = \|\rr\|$, and $\nn$ is the unit outward
normal, solves the incompressible Stokes equations for any sufficiently
smooth density function $\eeta$.  However, the double-layer potential can not capture velocity fields
corresponding to rigid body motions~\cite{pow-mir1987}.  A standard
method to complete the double-layer potential is to introduce a
Stokeslet and rotlet due to each interior body
\begin{align}
  S[\llambda_k,\cc_k] = \frac{1}{4\pi} \left(-\log \rho \II + 
    \frac{\rr \otimes \rr}{\rho^2} \right)\llambda_k
  \quad \text{and} \quad
  R[\xi_k,\cc_k] = \frac{\rr^\perp}{\rho^2}\xi_k,
  \label{eqn:stokeslet_rotlet}
\end{align}
where $\cc_k$ is a point in the interior of $\Omega_k$, $\rr = \xx -
\cc_k$, and $\rr^\perp = (r_2,-r_1)$.  Then, the solution of the Stokes
equations with a Dirichlet boundary condition $\ff$ can be written in
terms of the completed double-layer potential
\begin{align}
  \uu(\xx) = \DD[\eeta](\xx) + \sum_{k=1}^{M} S[\llambda_k,\cc_k](\xx) +
              \sum_{k=1}^{M} R[\xi_k,\cc_k](\xx), \quad \xx \in \Omega,
  \label{eqn:completedDLP}
\end{align}
where the density function, Stokeslets, and rotlets satisfy the
second-kind integral equation
\begin{subequations}
\label{eqn:completed_DLP}
\begin{align}
  \ff(\xx) &= -\frac{1}{2}\eeta(\xx) + \DD[\eeta](\xx) + 
      \sum_{k=1}^{M} S[\llambda_k,\cc_k](\xx) +
      \sum_{k=1}^{M} R[\xi_k,\cc_k](\xx), \quad \xx \in \Gamma, 
      \label{eqn:DLP}\\
  \llambda_k &= \frac{1}{2\pi}\int_{\Gamma_k} \eeta(\yy) \, ds_\yy, 
  \label{eqn:stokeslet} \\
  \xi_k &= \frac{1}{2\pi}\int_{\Gamma_k} \yy^\perp \cdot \eeta(\yy)
  \, ds_\yy.
  \label{eqn:rotlet}
\end{align}
\end{subequations}
Finally, because of the incompressibility constraint, the boundary
condition must satisfy 
\begin{align}
  \int_{\Gamma} \ff \cdot \nn \, ds_\yy = 0.
\end{align}
This results in a rank-one null space of
equation~\eqref{eqn:completed_DLP} which we remove by adding the 
rank-one modification $\NN_0[\eeta](\xx) = \int_{\Gamma_0} (\nn(\xx) \times
\nn(\yy))\nn(\yy) \, ds_\yy$, for $\xx \in \Gamma_0$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Computing the Shear Stress, Pressure, and Drag} 
\label{sec:qois}
The governing equations~\eqref{eqn:erosionModel} require the shear
stress which is defined on the boundary of each body.  The shear stress
is given by
\begin{align}
  \tau = \left(\nabla \uu + \nabla \uu^T \right)\nn \cdot \ss,
  \label{eqn:shearStress}
\end{align}
where $\ss$ is the unit tangent vector.  We note that for the Stokes
equations, because of the no-slip boundary condition on the bodies, the
shear stress is identical to the surface vorticity $\tau = u_y - v_x$.
However, we use equation~\eqref{eqn:shearStress} to compute the shear
stress since the jump conditions resulting from the double-layer
potential are known~\cite{qua-bir2014a}.  The velocity field of the
completed double-layer potential involves three terms: the double-layer
potential $\DD$, the Stokeslets $S$, and the rotlets $R$.  To compute
the shear stress, we first compute the deformation tensor $\ssigma =
\frac{1}{2}(\nabla\uu + \nabla\uu^T)$ of each of these terms
individually.

Considering the double-layer potential~\eqref{eqn:velocityDLP} evaluated
at $\xx \in \Omega$, the deformation tensor is
\begin{align}
  \ssigma^\DD(\xx) = \int_{\bd\Omega} K_\ssigma(\xx,\yy) \eeta(\yy) \,ds_\yy,
  \label{eqn:shearStressDLP}
\end{align}
where the kernel $K_\ssigma$ is given by
\begin{align*}
  K_\ssigma(\xx,\yy) = \frac{1}{2\pi}\left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) +
    \frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
  \right).
\end{align*}
We are interested in the deformation tensor at boundary points $\xx_0
\in \bd\Omega$, and this requires an expression for the jump in
$\ssigma^\DD(\xx)$ as $\xx \rightarrow \xx_0 \in \Gamma$.  The limiting
value is given by~\cite{qua-bir2014a}
\begin{align}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}\sigma_\DD(\xx) =
  J[\eeta](\xx_0) + \ssigma^\DD(\xx_0) = 
  \frac{1}{2}\pderiv{\eeta}{\ss} \cdot \ss
  \left[ \begin{array}{cc}
    s_x^2 - s_y^2 & 2s_x s_y \\ 2s_x s_y & s_y^2 - s_x^2
  \end{array}\right] + \ssigma^\DD(\xx_0), \quad \xx_0 \in \Gamma.
\end{align}

The deformation tensor due to the Stokeslets and rotlets in
equation~\eqref{eqn:stokeslet_rotlet} are
\begin{align}
  \ssigma^S(\xx_0) &= \frac{\rr \cdot \llambda}{4\pi\rho^2} \left(
    \II - \frac{2}{\rho^2} \rr \otimes \rr \right),  \\
  \ssigma^R(\xx_0) &= -\frac{\xi}{\rho^4} \left(\rr \otimes \rr^\perp + 
                           \rr^\perp \otimes \rr \right).
\end{align}
Now that the deformation tensor of each component of the completed
double-layer potential are computed, the shear stress at $\xx_0 \in
\Gamma$ can be computed as
\begin{align}
  \tau = \left(J[\eeta](\xx_0) + \ssigma^\DD(\xx_0) + 
    \ssigma^S(\xx_0) + \ssigma^R(\xx_0)\right) \nn \cdot \ss.
\end{align}

To compute the pressure at $\xx_0 \in \Gamma$, we follow the same
procedure used for the shear stress.  In particular, we compute the
pressure due to each of the three terms that compose the completed
double-layer potential.  They are given by
\begin{align}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}p^\DD(\xx) =  
  -\pderiv{\eeta}{\ss} \cdot \ss + p^{\DD}(\xx_0),
\end{align}
where
\begin{align}
  p^\DD(\xx) = -\frac{1}{\pi}\int_{\Gamma} \frac{1}{\rho^2}
    \left(I - 2 \frac{\rr \otimes \rr}{\rho^2}\right) 
    \nn \cdot \eeta(\yy) \,ds_\yy.
\end{align}
The pressure due to the Stokeslets is
\begin{align}
  p^S(\xx_0) = \frac{\rr \cdot \llambda}{2\pi\rho^2},
  \label{eqn:pressureDLP}
\end{align}
and the velocity due to a rotlet is pressure-free.  Therefore, once the
density function and Stokeslets are computed, the pressure at a point
$\xx_0 \in \Gamma$ is
\begin{align}
  p(\xx_0) = -\pderiv{\eeta}{\ss} \cdot \ss + p^{\DD}(\xx_0) + 
              \frac{\rr \cdot \llambda}{2\pi\rho^2}.
\end{align}

Finally, with the pressure and shear stress, the drag on a single body
with boundary $\Gamma_k$ is
\begin{align}
  F_D = \int_{\Gamma_k} (-p \nn + \tau \ss)\,ds.
\end{align}

\todo[inline]{Describe convention for normal and tangent vectors}

@@

For the velocity field, the kernel of the integral operator
\begin{align}
  D(\xx,\yy) = \frac{\rr \cdot \nn}{\rho^2} 
               \frac{\rr \otimes \rr}{\rho^2}
\end{align}
has a removable singularity when $\xx = \yy$.  In particular
\begin{align}
  \lim_{\substack{\yy \rightarrow \xx \\ \yy \in \Gamma}} D(\xx,\yy) = 
  \frac{\kappa(\xx)}{2\pi}\ss \otimes \ss,
\end{align}
where $\kappa(\xx)$ is the curvature of $\Gamma$ at $\xx$.  By using
this limiting term in when applying the trapezoid rule, spectral
accuracy is achieved because of the smoothness and periodicity of the
integrand~\cite{tre-wei2014}.

For the shear stress and the pressure, we will use odd-even integration
which achieves spectral accuracy if the integrand is no stronger than
$\mathcal{O}(\rho^{-1})$.  Since both these integrands are too strong,
the singularity must be weakened.  This can be done by adding a constant
density function which has no effect on the shear stress and
pressure~\cite{poz1992}.  In particular, the shear stress due to the
double-layer potential is equivalent to
\begin{align}
  \ssigma^{\DD}(\xx_0) &= \int_{\bd\Omega}K_\ssigma(\xx_0,\yy)
      \ssigma(\yy) \, ds_{\yy} \\ 
      &= \int_{\bd\Omega}K_\ssigma(\xx_0,\yy)
      (\ssigma(\yy) - \ssigma(\xx_0)) \, ds_{\yy}.
\end{align}
For the pressure, the double-layer potential is equivalent to
\begin{align}
  p^\DD(\xx) = -\frac{1}{\pi}\int_{\Gamma} \frac{1}{\rho^2}
    \left(I - 2 \frac{\rr \otimes \rr}{\rho^2}\right) 
    \nn \cdot (\eeta(\yy) - \eeta(\xx_0)) \,ds_\yy.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fast Summation} 
The majority of the computational cost comes from evaluating the the
double-layer potential $\mathcal{D}$ in equation~\eqref{eqn:DLP}.  This
layer potential can be efficiently evaluated using a fast summation
method such as the Fast Multipole Method
(FMM)~\cite{gre-rok1987}\footnote{The FMM for the Stokes double-layer
poetntial was supplied by Manas Rachh.}.  By using the
FMM,~\eqref{eqn:completed_DLP} can be solved with $\mathcal{O}(N)$
operations, where $N$ is the total number of discretization points.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Formulation\label{s:formulation}} 
Let us first define the main variables used to model erosion.  We only
consider flows that are confined by a solid wall $\Gamma$ that encloses
$M$ eroding bodies.  The bodies are denoted as $\gamma_k$,
$k=1,\ldots,M$, and we write $\gamma = \gamma_1 \cup \cdots \cup
\gamma_M$.  Neglecting inertial forces, the dynamics of the fluid is
fully characterized by the position of the bodies $\xx_k(s,t) \in
\gamma_k$, where $s$ is the arclength and $t$ is time.  Given
$\xx_k(s,t)$, $k=1,\ldots,M$, derived variables include the fluid
velocity $\uu$, the pressure $p$, and the shear stress $\tau$.   On the
bounding wall, we prescribed a velocity $\UU(\xx,t)$.  Then, the
governing equations are
\begin{equation}
\label{eqn:erosionModel}
\begin{split}
  \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap &&\mbox{conservation
of momentum}\\
\grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap
&&\mbox{conservation of mass} \\
\uu = 0, &\hspace{20pt} \xx \in \gamma, \gap &&\mbox{no slip on the
bodies} \\
\uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap &&\mbox{outer wall
velocity} \\
\dot{\xx}(t) = -|\tau| \nn, &\hspace{20pt} \xx \in \gamma,
&&\mbox{erosion model},
\end{split}
\end{equation}
where $\nn$ is the outward unit normal vector which points away from the
fluid.  The outer wall is formed by rounding off the corners of $[-3,3]
\times [-1,1]$, and we enforce the Hagen-Poiseuille flow
\begin{align*}
  \uu(\xx) = (1-y^2,0)
\end{align*}
on $\Gamma$.  The boundary of the fluid domain is $\bd\Omega = \gamma_1
\cup \cdots \cup \gamma_M \cup \Gamma$.  To minimize the boundary
effects of the inflow and outflow, we only place bodies in the center
third, $[-1,1] \times [-1,1]$, of the fluid domain.
\begin{figure}[htpb]
  \centering
  \input{figs/schematic.tikz}
  \caption{\label{fig:schematic} A schematic for the governing
    equations.  A no-slip boundary condition is imposed on each body
    $\gamma_k$ whose unit normal points outward relative to the
    geometry.  On the outer geometry $\Gamma$, a pipe flow is imposed.
    Finally, the bodies are eroded according to the shear stress applied
    by the Stokesian fluid.}
\end{figure}

We use a quasi-static approximation.  That is, at each time step, we
freeze the geometry, solve the incompressible Stokes equations, and then
update the geometry according to the erosion model.  A flow diagram is
illustrated in Figure~\ref{fig:workflow}.  The two steps inside the red
rectangle describes a time step in the quasi-static approximation
\begin{figure}[htpb]
  \centering
  \input{figs/workflow.tikz}
  \caption{\label{fig:workflow}A caption.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Boundary Integral Equation Formulation} 
\label{sec:bies}
Here we focus on solving the incompressible Stokes equations with
Dirichlet boundary conditions
\begin{alignat*}{3}
  \mu\Delta \uu &= \grad p, \qquad && \xx \in \Omega, \\
  \grad \cdot \uu &= 0,   && \xx \in \Omega, \\
  \uu &= \ff,  && \xx \in \bd\Omega.
\end{alignat*}
The viscosity $\mu$ sets the time scale, and we will assume that
$\mu=1$.  We start by defining the double-layer potential, which is the
convolution of the stresslet with a function $\eeta:\bd\Omega
\rightarrow \RR^2$,
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{\pi} \int_{\Gamma} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2} 
    \eeta \, ds_\yy, \quad \xx \in \Omega,
    \label{eqn:velocityDLP}
\end{align}
where $\rr = \xx - \yy$, $\rho = \|\rr\|$, and $\nn$ is the unit outward
normal.  For any density function $\eeta$, the double-layer potential
satisfies the incompressible Stokes equations, but the double-layer
potential can not capture velocity fields corresponding to rigid body
motions~\cite{pow-mir1987}.  A standard method to complete the
double-layer potential is to introduce a Stokeslet and rotlet due to
each interior body
\begin{align}
  S[\llambda_k,\cc_k] = \frac{1}{4\pi} \left(-\log \rho \II + 
    \frac{\rr \otimes \rr}{\rho^2} \right)\llambda_k
  \quad \text{and} \quad
  R[\xi_k,\cc_k] = \frac{\rr^\perp}{\rho^2}\xi_k,
  \label{eqn:stokeslet_rotlet}
\end{align}
where $\cc_k$ is a point inside the $k^{th}$ body,  $\rr = \xx - \cc_k$,
and $\rr^\perp = (r_2,-r_1)$.  Then, the solution of the Stokes
equations with a Dirichlet boundary condition $\ff$ is
\begin{align}
  \label{eqn:completed_DLP}
  \uu(\xx) = \DD[\eeta](\xx) + \sum_{k=1}^{M} S[\llambda_k,\cc_k](\xx) +
              \sum_{k=1}^{M} R[\xi_k,\cc_k](\xx), \quad \xx \in \Omega,
\end{align}
where the density function, Stokeslets, and rotlets satisfy
\begin{subequations}
\label{eqn:completed_BIE}
\begin{align}
  \ff(\xx) &= -\frac{1}{2}\eeta(\xx) + \DD[\eeta](\xx) + 
      \sum_{k=1}^{M} S[\llambda_k,\cc_k](\xx) +
      \sum_{k=1}^{M} R[\xi_k,\cc_k](\xx), \quad \xx \in \Gamma, 
      \label{eqn:DLP}\\
  \llambda_k &= \frac{1}{2\pi}\int_{\Gamma_k} \eeta(\yy) \, ds_\yy, 
  \label{eqn:stokeslet} \\
  \xi_k &= \frac{1}{2\pi}\int_{\Gamma_k} \yy^\perp \cdot \eeta(\yy)
  \, ds_\yy.
  \label{eqn:rotlet}
\end{align}
\end{subequations}
Finally, because of the incompressibility constraint, the boundary
condition must satisfy 
\begin{align*}
  \int_{\Gamma} \ff \cdot \nn \, ds_\yy = 0.
\end{align*}
The result is a rank-one null space of~\eqref{eqn:completed_BIE} which
is removed by adding the rank-one modification $\NN_0[\eeta](\xx) =
\int_{\Gamma_0} (\nn(\xx) \otimes \nn(\yy))\nn(\yy) \, ds_\yy$, for $\xx
\in \Gamma_0$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Computing the Shear Stress, Pressure, and Drag} 
\label{sec:qois}
There are several quantities besides the velocity that need to be
comnputed.  First, the shear stress along $\gamma_k$ is required by the
governing equations~\eqref{eqn:erosionModel}.  In addition, to study
characteristics of the flow, the pressure and drag must also be
computed.  Once equation~\eqref{eqn:completed_DLP} is solved for the
density function, Stokeslets, and rotlets, all of these quantities can
be computed by evaluating layer potentials that we now derive.

\subsubsection{Shear Stress}
The shear stress is defined by
\begin{align}
  \tau = \left(\nabla \uu + \nabla \uu^T \right)\nn \cdot \ss,
  \label{eqn:shearStress}
\end{align}
where $\ss$ is the unit tangent vector (Figure~\ref{fig:schematic}).  We
note that for the Stokes equations, because of the no-slip boundary
condition on the bodies, the shear stress is identical to the surface
vorticity $\tau = u_y - v_x$.  However, we use
equation~\eqref{eqn:shearStress} to compute the shear stress since the
jump conditions resulting from the double-layer potential are
known~\cite{qua-bir2014a}.  The velocity field of the completed
double-layer potential~\eqref{eqn:completed_DLP} involves three terms:
the double-layer potential $\DD$, the Stokeslets $S$, and the rotlets
$R$.  To compute the shear stress, we first compute the deformation
tensor $\ssigma = \frac{1}{2}(\nabla\uu + \nabla\uu^T)$ of each of these
terms individually.

The deformation tensor of the double-layer
potential~\eqref{eqn:velocityDLP} at $\xx \in \Omega$ is
\begin{equation}
  \label{eqn:shearStressDLP}
  \begin{aligned}
  \ssigma^\DD(\xx) = \frac{1}{2\pi}\int_{\bd\Omega} &\left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) 
    \right. \\
    &\left.
    +\frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
  \right) ds_\yy.
  \end{aligned}
\end{equation}
We are interested in the deformation tensor at boundary points $\xx_0
\in \bd\Omega$, and this requires an expression for the jump in
$\ssigma^\DD(\xx)$ as $\xx \rightarrow \xx_0 \in \Gamma$.  The limiting
value is given by~\cite{qua-bir2014a}
\begin{align*}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}\sigma_\DD(\xx) =
  J[\eeta](\xx_0) + \ssigma^\DD(\xx_0) = 
  \frac{1}{2}\pderiv{\eeta}{\ss} \cdot \ss
  \left[ \begin{array}{cc}
    s_x^2 - s_y^2 & 2s_x s_y \\ 2s_x s_y & s_y^2 - s_x^2
  \end{array}\right] + \ssigma^\DD(\xx_0), \quad \xx_0 \in \Gamma.
\end{align*}

The deformation tensor due to the Stokeslets and
rotlets~\eqref{eqn:stokeslet_rotlet} are
\begin{align*}
  \ssigma^S(\xx_0) &= \frac{\rr \cdot \llambda}{4\pi\rho^2} \left(
    \II - \frac{2}{\rho^2} \rr \otimes \rr \right),  \\
  \ssigma^R(\xx_0) &= -\frac{\xi}{\rho^4} \left(\rr \otimes \rr^\perp + 
                           \rr^\perp \otimes \rr \right).
\end{align*}
Now that the deformation tensor of each component of the completed
double-layer potential are computed, the shear stress at $\xx_0 \in
\Gamma$ can be written as
\begin{align*}
  \tau = \left(J[\eeta](\xx_0) + \ssigma^\DD(\xx_0) + 
    \ssigma^S(\xx_0) + \ssigma^R(\xx_0)\right) \nn \cdot \ss.
\end{align*}

\subsubsection{Pressure}
To compute the pressure at $\xx_0 \in \Gamma$, we follow the same
procedure used for the shear stress.  In particular, we compute the
pressure due to each of the three terms in the completed double-layer
potential~\eqref{eqn:completed_DLP}.  They are given by
\begin{align*}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}p^\DD(\xx) =  
  -\pderiv{\eeta}{\ss} \cdot \ss + p^{\DD}(\xx_0),
\end{align*}
where
\begin{align*}
  p^\DD(\xx) = -\frac{1}{\pi}\int_{\Gamma} \frac{1}{\rho^2}
    \left(I - 2 \frac{\rr \otimes \rr}{\rho^2}\right) 
    \nn \cdot \eeta(\yy) \,ds_\yy.
\end{align*}
The pressure due to the Stokeslets is
\begin{align}
  p^S(\xx_0) = \frac{\rr \cdot \llambda}{2\pi\rho^2},
  \label{eqn:pressureDLP}
\end{align}
and the pressure of a rotlet is zero.  Therefore, once the density
function and Stokeslets are computed, the pressure at a point $\xx_0 \in
\Gamma$ is
\begin{align*}
  p(\xx_0) = -\pderiv{\eeta}{\ss} \cdot \ss + p^{\DD}(\xx_0) + 
              \frac{\rr \cdot \llambda}{2\pi\rho^2}.
\end{align*}

\subsubsection{Drag}
Finally, with the pressure and shear stress, the drag on a single body
with boundary $\Gamma_k$ is
\begin{align*}
  F_D = \int_{\Gamma_k} (-p \nn + \tau \ss)\,ds.
\end{align*}

@@

For the velocity field, the kernel of the integral operator
\begin{align*}
  D(\xx,\yy) = \frac{\rr \cdot \nn}{\rho^2} 
               \frac{\rr \otimes \rr}{\rho^2}
\end{align*}
has a removable singularity when $\xx = \yy$.  In particular
\begin{align*}
  \lim_{\substack{\yy \rightarrow \xx \\ \yy \in \Gamma}} D(\xx,\yy) = 
  \frac{\kappa(\xx)}{2\pi}\ss \otimes \ss,
\end{align*}
where $\kappa(\xx)$ is the curvature of $\Gamma$ at $\xx$.  By using
this limiting term in when applying the trapezoid rule, spectral
accuracy is achieved because of the smoothness and periodicity of the
integrand~\cite{tre-wei2014}.

For the shear stress and the pressure, we will use odd-even integration
which achieves spectral accuracy if the integrand is no stronger than
$\mathcal{O}(\rho^{-1})$.  Since both these integrands are too strong,
the singularity must be weakened.  This can be done by adding a constant
density function which has no effect on the shear stress and
pressure~\cite{poz1992}.  In particular, the shear stress due to the
double-layer potential is equivalent to
\begin{align*}
  \ssigma^{\DD}(\xx_0) &= \int_{\bd\Omega}K_\ssigma(\xx_0,\yy)
      \ssigma(\yy) \, ds_{\yy} \\ 
      &= \int_{\bd\Omega}K_\ssigma(\xx_0,\yy)
      (\ssigma(\yy) - \ssigma(\xx_0)) \, ds_{\yy}.
\end{align*}
For the pressure, the double-layer potential is equivalent to
\begin{align*}
  p^\DD(\xx) = -\frac{1}{\pi}\int_{\Gamma} \frac{1}{\rho^2}
    \left(I - 2 \frac{\rr \otimes \rr}{\rho^2}\right) 
    \nn \cdot (\eeta(\yy) - \eeta(\xx_0)) \,ds_\yy.
\end{align*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fast Summation} 
The majority of the computational cost comes from evaluating the the
double-layer potential $\mathcal{D}$ in equation~\eqref{eqn:DLP}.  This
layer potential can be efficiently evaluated using a fast summation
method such as the Fast Multipole Method
(FMM)~\cite{gre-rok1987}\footnote{The FMM for the Stokes double-layer
poetntial was supplied by Manas Rachh.}.  By using the
FMM,~\eqref{eqn:completed_BIE} can be solved with $\mathcal{O}(N)$
operations, where $N$ is the total number of discretization points.
